Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-10-13 algorithm.tex
AA: fornendo motivazioni alle metriche scelte
AA: altre modifiche

	\begin{enumerate}
	\item \emph{high innovation regions}, whose innovation score exceeds the average score of the previous era by a factor of $\alpha$, i.e.
	\[
	\mathcal{R}_{i,h}=\left\{ \left.R\in\mathcal{R}_{i}\right|is\left(R\right)>\alpha\cdot is_{av,i-1}\right\} 
	\]
	 where $\alpha$ is a constant defined by the designer (for example
	$\alpha=1.2$). 
	\item \emph{low innovation regions}, whose score is positive but modest, i.e.
	\[
	\mathcal{R}_{i,l}=\left\{ \left.R\in\mathcal{R}_{i}\right|0<is\left(R\right)\le\alpha\cdot is_{av,i-1}\right\} 
	\]
	\item \emph{no innovation regions}, whose innovation is null, i.e. 
	\[
	\mathcal{R}_{i,n}=\left\{ \left.R\in\mathcal{R}_{i}\right|is\left(R\right)=0\right\} 
	\]
	\end{enumerate}

@@

>>>>

	\begin{enumerate}
	\item \emph{high innovation regions}, whose innovation score exceeds the average score of the previous era by a factor of $\alpha$, i.e.
	\[
	\mathcal{R}_{i,h}=\left\{ \left.R\in\mathcal{R}_{i}\right|is\left(R\right)>\alpha\cdot is_{av,i-1}\right\} 
	\]
	 where $\alpha$ is a constant defined by the designer (for example
	$\alpha=1.2$). 	
		\comment{AA: Secondo me e' strano usare $\alpha\cdot is_{av,i-1}$ come soglia. Non si capisce perche' non si usi $is_{av,i}$ ma si usi quello dell'era precedente. Non e' motivata da nessuna parte. Non si capisce perche', in una finale di miss italia, io dovrei conforntare le ragazze di oggi con la ragazza media dell'anno scorso. E' piu' naturale confrontare le ragazze di oggi con la ragazza media di oggi stesso, per vedere chi e' oltre alla media attuale. Che ne dici di cambiare l'espressione in 
			\[
			\mathcal{R}_{i,h}=\left\{ \left.R\in\mathcal{R}_{i}\right|is\left(R\right)>\alpha\cdot is_{av,i}\right\} 
			\]
		}
	\item \emph{low innovation regions}, whose score is positive but modest, i.e.
	\[
	\mathcal{R}_{i,l}=\left\{ \left.R\in\mathcal{R}_{i}\right|0<is\left(R\right)\le\alpha\cdot is_{av,i-1}\right\} 
	\]
	\item \emph{no innovation regions}, whose innovation is null, i.e. 
	\[
	\mathcal{R}_{i,n}=\left\{ \left.R\in\mathcal{R}_{i}\right|is\left(R\right)=0\right\} 
	\]
	\end{enumerate}

@@

\subsection{Algorithm parameters}
The parameters that completely determine the algorithm are $K$ and $\alpha$.
$K$ corresponds to the population size of the genetic approach. Authors of \cite{zitzler_ec00} found that increasing $k$ not necessarily has a considerable impact on the performance of the algorithm. This holds also in our algorithm.  In addition, while in the genetic approach $K$ must be sufficiently large in order to guarantee diversity of the Pareto set configurations, in our algorithm the diversity is not impacted by $K$ since, by construction, it is guaranteed by the way the configuration space is partitioned, independently from number of configurations evaluated at each iteration. \comment{This is confirmed by the experimental results in \secR{parameter_impact}. }

Parameter $\alpha$ determines when a region is considered of high innovation. High values of $\alpha$ mean that a region will hardly be elected as high-innovation, yielding less splitting operations. This slows down zooming in on portions of configuration space and make the distribution of the simulated configuration more uniform, as there are fewer small regions with a high test configuration concentration. The opposite happens when small values of $\alpha$ are used. \comment{This is confirmed by results of \secR{parameter_impact}. }

### 2014-10-13 appendix.tex
AA: altre modifiche

Metto qua tutte le cose che non si possono lasciare nel testo ma che e' un peccato buttare nel dimenticatoio

\begin{definition}[Separation (alternative)] Another definition of separation could be:
\[
s\left(\mathbf{o}\rightarrow\mathbf{q}\right)=\max_{i=1,\dots,m}\left|\frac{o_{i}-q_{i}}{o_{i}}\right|
\]

\end{definition}

>>>>

\section{appendix}
Metto qua tutte le cose che non si possono lasciare nel testo ma che e' un peccato buttare nel dimenticatoio

\begin{definition}[Separation (alternative)] Another definition of separation could be:
	\[
	s\left(\mathbf{o}\rightarrow\mathbf{q}\right)=\max_{i=1,\dots,m}\left|\frac{o_{i}-q_{i}}{o_{i}}\right|
	\]
\end{definition}

\subsection{Virtual merging}
PROBLEMA: Secondo il punto debole del nostro algoritmo, che fa si che noi ci facciamo sempre fregare da GA e' il merging. Il merging puo' essere effettuato solo tra regioni contigue. Non so come siano andati i run, ma ad intuito, trovare delle no-innovation-regions che sono pure contigue e' proprio una botta di fortuna. Secondo me la maggior parte delle no-innovation-regions sono non contigue e quindi non si mergiano, continuando a restistere imperterritamente col procedere dei run. Risultato: le no-innovation-regions continuano a fregarsi, ad ogni iterazione, piu' run di quanti dovrebbero.

SOLUZIONE: Forse per migliorare l'algoritmo si dovrebbe abbandonare il concetto di merging, visto che mi sa che per ora e' sulla carta, ma durante i run avviene poche volte. In alternativa, si potrebbe usare un fattore di qualita', che simuli l'avvenimento dei merge. Ad ogni regione $R$ si associa un fattore di qualita' $q_R$, di default 0. Quando la regione e' classificata come no-innovation, $q_R$ si decrementa. A ciascuna iterazione, anziche' assegnare ad ogni regione K/N simulazioni, dove N è il numero di regioni, ogni regione $R*$ dovrebbe avere il seguente numero di simulazioni
	\begin{align}
    \frac{2^{q_R*} K} { \sum_R  2^{q_R} }
	\end{align}

\subsection{Importance given to the innovation score}
AND: una prova che si potrebbe fare e' modificare leggerment l'innovation score di una regione. Forse questa metrica da troppo peso all'innovation delle configurazioni. Se una regione e' ricca di pareto points, ma essi sono tutti simili (nel senso che le loro immagini nello spazio degli obiettivi sono tutte vicine), dopo un po' di iterazioni l'analisi di quella regione non portera' sufficiente innovazione e la regione sara' abbandonata. Il che e' un peccato, perche' magari quei punti che in questo modo ci stiamo perdendo permetterebbero di fregare, o eguagliare, genetic algo. Questo spiega anche perche' le nostre soluzioni spaziano di piu' lo spazio degli obiettivi e vanno piu' verso gli estremi. Se volessimo rinunciare a spaziare per forza tutto il range degli obiettivi, volendo essere piu' simili e competitivi con genetic algorithm, potremmo modificare il calcolo dell'innovation score (step (4) della sezione 4.2), nel seguente modo:
	\begin{align}
	    is(c^*) = 1 + B\cdot \min[ s(c \rightarrow c^*) | c \in \mathscr{P}_{i-1} ]
	\end{align}
In questo modo, il fattore B permette di pesare quanta importanza dare all'innovativita' del nuovo pareto point. Ad es, per B=0, tutti i nuovi pareto points sono buoni uguali, a prescindere dalla loro innovativita'. Per B=0.2, un nuovo pareto point gia' e' una cosa bella, se poi ha innovativita', meglio ancora, sarebbe la ciliegina sulla torta, ecc .... Si potrebbero fare nuovi run con questo calcolo di innovation score, io li farei per B=1 e B=0.5, per vedere se le cose migliorano.

### 2014-10-13 experiments.tex
AA: altre modifiche

% ACM table style
\begin{table}
	\tbl{Benchmarks\label{tab:bench}}{
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Benchmark} & \multicolumn{1}{c}{Application} \\
	\hline
	Alloca\_test & Memory array allocation test \\
	Bmm & Matrices multiplication and elements sum \\
	Fir\_int & Finite impulse response \\
	Mm & Floating point matrices multiplication \\
	Sha256 & Cryptocurrency header Hashing \\
	Wave & Wavefront computation \\
	\hline
	\end{tabular}}
\end{table}
\subsection{Terms of comparison}

@@

@@

\subsection{Evaluation metrics}
In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}.

For a given objective, the \emph{variation range} represents the ratio between the maximum and the minimum value observed for that objective.
\comment{AA: Per rappresentare l'extent del Pereto front, forse sarebbe meglio evitare di usare il variation range, per come l'hai definito tu, e usare la funzione M3 di \cite{zitzler_ec00}, Sec.5 ``Metrics of performance'', che mi pare rappresenti la stessa cosa, concettualmente. Questo ci permette di essere meno attaccabili (del tipo ``non te la prendere con noi, prenditela con Zitzler se non capisci l'utilita' della metrica'').}
It quantifies the extent of the of the Pareto front that is an important factor to consider when evaluating the quality of a Pareto front. Indeed, authors in \cite{zitzler_ec00} carefully observe that an exploration algorithm should preferably provide pareto fronts whose objective function range is large.
Such Pareto fronts permit to investigate in a broader way the potentials of the designed device. This is important, in general, and turns out to be essential when designing a device that may be required to work in different operational conditions, for example with stringent execution time constraints and less strict power dissipation ones or, on the contrary, with low power requirements and tolerance to high execution time.

The average \emph{normalised absolute dispersion error} measures
the average absolute difference between the distribution of points in
the objective space and an ideal distribution in which the points are
uniformly distributed over the objective space. 
The motivation behind this metric is, again, given by \cite{zitzler_ec00}, in which authors remark that a desirable property of the Pareto front is the spread of the non-dominated configurations all over the configuration space. In other words, the more similar the non-dominated configuration distribution is to the uniform distribution, the better is the quality of the Pareto front.
\comment{AA: commentare perche' e' importante questa metrica. Cercare la parola chiave ``deception'' o qualcosa del genere}.
Formally, let $O$ be
the image, in the objective space, of the configurations visited by
the design space exploration. The generic element of $O$ (\ie, a
solution) is a pair $(p,t)$ where $p$ and $t$ are the average power
and execution time, respectively. The two-dimensional objective space
is then partitioned by a $M_x \times M_y$ mesh. For each tile $T_i,
\ i=1, 2, \ldots, M_xM_y$ of the mesh, let $N_i$ be the number
of points in $O$ which fall in $T_i$. The average absolute error, $E_i$, for
$T_i$ is the absolute value of the difference between $N_i$ and the
ideal number of solutions, $\overline{N}$, which should fall in $T_i$
in case of uniform distribution. Such $\overline{N}$ can be simply
computed as the ratio between the cardinality of $O$ and the number of
tiles. Thus,
\[ E_i = |N_i - \overline{N}|, \]
where $\overline{N} = |O| / (M_x M_y)$. The average
normalised absolute dispersion error ($ANADE$) is the average of $E_i$
normalised to the maximum absolute error $E_{max}$:
\[ ANADE = \frac{\sum_{i=1}^{M_xM_y} E_i/(M_xM_y)}{E_{max}}, \]
where $E_{max}$ can be computed as the average absolute error in the worst
case in which all the solutions fall in a single tile:
\[ E_{max} = \frac{(M_x M_y - 1) \overline{N} + |\overline{N} -
    |O|| }{M_x M_y}. \]

@@

\figR{pareto_fronts_100} and \figR{pareto_fronts_500} show the resulting Pareto fronts for two
different scenarios of 100 and 500 simulations budget. When possible,
similar algorithm parameters have been adopted. In particular a
population size of 30 has been chosen for the genetic approach (GA), corresponding to the
same value for the $K$ of PS. The number of generations (or era)
has been chosen so that the only the given amount of simulations is
actually performed. Thus, when accounting these simulations only unique simulations
has been considered, removing repetitions and unfeasible
configurations.

>>>>

% ACM table style
\begin{table}
	\tbl{Benchmarks\label{tab:bench}}{
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Benchmark} & \multicolumn{1}{c}{Application} \\
	\hline
	Alloca\_test & Memory array allocation test \\
	Bmm & Matrices multiplication and elements sum \\
	Fir\_int & Finite impulse response \\
	Mm & Floating point matrices multiplication \\
	Sha256 & Cryptocurrency header Hashing \\
	Wave & Wavefront computation \\
	\hline
	\end{tabular}}
\end{table}

\subsection{Terms of comparison}

@@

%------------------------------------------------------------------------------
\subsection{Evaluation campaign}
In order to obtain the results presented in this section, we implemented our exploration strategy in Epic Explorer~\cite{ascia2003epic}, an open source framework for the design space explora=tion of embedded systems. \comment{AA: continuare descrivendo un po' il workflow. The reader can refer to~\cite{ascia2003epic} for more details about the estimation models of the power dissipation and execution time. Scrivere che it is possible to run benchmarks }.

@@

\subsection{Evaluation metrics}
In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. Both of them permit to quantify the extent of the Pareto front that is an important factor to consider when evaluating the quality of a Pareto front. Indeed, authors in \cite{zitzler_ec00} carefully observe that an exploration algorithm should preferably provide pareto fronts whose objective function range is large.
Such Pareto fronts permit to investigate in a broader way the potentials of the designed device. This is important, in general, and turns out to be essential when designing a device that may be required to work in different operational conditions, for example with stringent execution time constraints and less strict power dissipation ones or, on the contrary, with low power requirements and tolerance to high execution time.

For a given objective, the \emph{variation range} represents the ratio between the maximum and the minimum value observed for that objective.
\comment{AA: Per rappresentare l'extent del Pereto front, forse sarebbe meglio evitare di usare il variation range, per come l'hai definito tu, e usare la funzione M3 di \cite{zitzler_ec00}, Sec.5 ``Metrics of performance'', che mi pare rappresenti la stessa cosa, concettualmente. Questo ci permette di essere meno attaccabili (del tipo ``non te la prendere con noi, prenditela con Zitzler (che e' citato 3064 volte!) se non capisci l'utilita' della metrica'').}

The average \emph{normalised absolute dispersion error} measures
the average absolute difference between the distribution of points in
the objective space and an ideal distribution in which the points are
uniformly distributed over the objective space. 
Formally, let $O$ be
the image, in the objective space, of the configurations visited by
the design space exploration. The generic element of $O$ (\ie, a
solution) is a pair $(p,t)$ where $p$ and $t$ are the average power
and execution time, respectively. The two-dimensional objective space
is then partitioned by a $M_x \times M_y$ mesh. For each tile $T_i,
\ i=1, 2, \ldots, M_xM_y$ of the mesh, let $N_i$ be the number
of points in $O$ which fall in $T_i$. The average absolute error, $E_i$, for
$T_i$ is the absolute value of the difference between $N_i$ and the
ideal number of solutions, $\overline{N}$, which should fall in $T_i$
in case of uniform distribution. Such $\overline{N}$ can be simply
computed as the ratio between the cardinality of $O$ and the number of
tiles. Thus,
\[ E_i = |N_i - \overline{N}|, \]
where $\overline{N} = |O| / (M_x M_y)$. The average
normalised absolute dispersion error ($ANADE$) is the average of $E_i$
normalised to the maximum absolute error $E_{max}$:
\[ ANADE = \frac{\sum_{i=1}^{M_xM_y} E_i/(M_xM_y)}{E_{max}}, \]
where $E_{max}$ can be computed as the average absolute error in the worst
case in which all the solutions fall in a single tile:
\[ E_{max} = \frac{(M_x M_y - 1) \overline{N} + |\overline{N} -
    |O|| }{M_x M_y}. \]

\comment{AA: Un'altra interessante metrica potrebbe essere questa: Auhtors of \cite{zitzler_ec00} remark that a desirable property of the Pareto front is the spread of the non-dominated configurations all over the configuration space. In other words, the more similar the non-dominated configuration distribution is to the uniform distribution, the better is the quality of the Pareto front, since we have not been captured by only one possible subset of Pareto configurations but we are able to span all over the configuration space and find configurations whose parameter values differ considerably. Mi aspetto che PS performi bene sotto questo punto di vista. Ovviamente, non so se e' il caso di considerarla ora. Se non e' il caso, puoi spostare questo testo nel file ``appendix.tex'' come possibile spunto futuro. Per quantificare questa cosa, si potrebbe usare come metrica la normalised absolute dispersion error pero' non nello spazio degli obiettivi, ma nello spazio delle configurazioni.}.

\subsection{Impact of the algorithm parameters}
\secL{parameter_impact}
\comment{IMPACT OF $K$. Si potrebbero presentare un plot in cui il numero totale di simulazioni e' fissato e vengono eseguiti diversi run dell'algoritmo con diversi valori di K. Per ogni run, il pareto set e' rappresentato nel plot, sottoforma di una curva diversa. Mi aspetterei che tutte le curve siano molto simili. Per ciascun run, il valore di variation range e average absolute dispersion error sono riportati. Mi aspetto valori simili, il che confermerebbe che la diversity nel nostro algoritmo non dipende da $K$. Ne vale la pena rifare questi run? Ci vuole troppo tempo?}

\comment{Si potrebbe mettere un plot con nelle ascisse l'era e nelle ordinate il numero di split avvenuti in quell'era. Si potrebbero disegnare diverse curve, una per ogni valore di alpha, per far vedere che il numero di split e' piu' alto q. Recommended values are close to 1, as results in \secR{ee} confirm. Using different values, if sufficiently close to 1, does not have a considerable impact on the resulting Pareto front approximation. Ne vale la pena rifare questi run? Ci vuole troppo tempo?}

@@

\figR{pareto_fronts_100} and \figR{pareto_fronts_500} show the resulting Pareto fronts for two
different scenarios of 100 and 500 simulations budget. \comment{AA: per curiosita': quando facevi piu' run, PS andava ancora peggio rispetto a GA?} When possible,
similar algorithm parameters have been adopted. In particular a
population size of 30 has been chosen for the genetic approach (GA), corresponding to the
same value for the $K$ of PS. \comment{AA: che valore di $alpha$ hai usato?} The number of generations (or era)
has been chosen so that the only the given amount of simulations is
actually performed. Thus, when accounting these simulations only unique simulations
has been considered, removing repetitions and unfeasible
configurations.
